---
title: "ChamberProcWorkflow"
output:
 rmarkdown::html_vignette:
    keep_md: true
vignette: >
  %\VignetteIndexEntry{ChamberProcWorkflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette explains how to use the package ChamberProc for Picarro data. The package largely depends on the package "RespChamberProc" and parts of the vignette are containing information from the "RespChamberProc" vignette.

First, the necessary libraries are loaded:

```{r setup, message=F}
# load libraries
library(ChamberProc)
library(RespChamberProc)
library(rlang)
library(changepoint)
library(nlme)
library(segmented)
library(tibble)
library(plyr)
library(dplyr)
library(purrr)
library(stringr)
library(ggplot2)
library(data.table)
library(lubridate)
library(tidyr)
library(doSNOW)
library(furrr)
library(elevatr)
library(sf)
library(openmeteo) #retrieves historical weather data from https://open-meteo.com/en/docs/historical-weather-api

```

# Preparing the chamber data

The user has to set the geographical coordinates of the study site (latitude and longitude) in degrees. Also, the Volue and surface Area of the Chamber need to be defined.

```{r user-defined input}
latDeciDeg <- 37.91514875822371 #enter here latitude in Geographical Coordinates (decimal degrees); crs =4326)
lonDeciDeg <- -4.72405536319233 #enter here longitude in Decimal Degrees

chamberVol = 0.6*0.6*0.6		
surfaceArea = 0.6*0.6

#path to results folder
dir.create("results")
results_dir <- "results/"
  
```

The data is loaded. This is a typical data.frame as obtained from a logger file of a PICARRO chamber. Sensor readings from various gases are recorded for sometimes non-equidistant timestamps <1 sec. In the following example, the sample dataset "sample_PICARRO_data" that comes with the "ChamberProc" package is used.

Sometimes, not the entire time series that was loaded is needed, therefore the data can be subset. Columns that are not needed for further calculations should be excluded.

```{r load-data}
ds0 <- sample_PICARRO_data

#create timestamp from date and time columns
ds0$TIMESTAMP <- as.POSIXct(paste0(ds0$DATE," ",ds0$TIME), "%Y-%m-%d %H:%M:%S", tz= "UTC")

# Subset data
ds_subset <- subset(ds0, as.numeric(TIMESTAMP) >= as.numeric(RespChamberProc::as.POSIXctUTC("2022-05-31 08:00:00")) )
ds_subset <- subset(ds_subset, as.numeric(TIMESTAMP) <= as.numeric(RespChamberProc::as.POSIXctUTC("2022-05-31 09:30:00" )) )

ds <- ds_subset %>% select(.,-c(DATE,TIME,FRAC_DAYS_SINCE_JAN1,FRAC_HRS_SINCE_JAN1,JULIAN_DAYS,EPOCH_TIME,ALARM_STATUS,INST_STATUS,CHAMBER_TEMP_sync,CHAMBER_PRESSURE_sync,SWITCH_sync,SOIL_TEMP_sync))

head(ds)
```

# Correct Gases

The PICARRO IRGA gives dry mole fractions for CO2, N2O, CH4, but not for NH3 and H2O. Therefore these gases have to be corrected. Additional envrionmental data (h, Pa and AirTemp are crucial) is loaded with the function "Additional_Weather_Data", which downloads meteorological data from the www.open-meteo.com API via the package "openmeteo" and elevation data from the "elevatR" package. The weather data comes in an hourly interval.
Alternatively, values for `r ds$h` and `r ds$AirTemp` ,and `r ds$AirTemp` can be assigned manually.

```{r, message=FALSE}

Additional_Weather_Data <- getAdditionalWeatherVariables(latDeciDeg, lonDeciDeg,format(min(ds$TIMESTAMP),"%Y-%m-%d"),format(max(ds$TIMESTAMP),"%Y-%m-%d"))

# For the case of PICARRO IRGA gives dry mole fractions for CO2, N2O, CH4, but not for NH3 and H2O
ds$solenoid_valvesInt<-  ds$solenoid_valves %>% as.integer()
ds$H2Oppt <- ds$H2O*10 # H2O from PICARRO is in %.Needs to be in ppt --> We need to multiply by 10
ds$N2O_dry <- ds$N2O_dry1min
ds$NH3_dry <- 10E-3*corrConcDilution(ds, colConc = "NH3", colVapour = "H2Oppt")  #NH3 from PICARRO is in ppb --> multiply colVapour by 10^-3 to get ppm
ds$H2O_dry <- corrConcDilution(ds, colConc = "H2Oppt", colVapour = "H2Oppt")

## (h, Pa, and AirTemp are obtained directly from freely available Meteo and Elevation data for given coordinates during the measurement time interval with the "getElevationAndAtmosphericPressure" script (see details therein). If desired, it can also be set here manually)
# ds$h = 106 # Córdoba elevation (m above sea level)
# ds$Pa <- 101325*(1 - 2.25577*10^-5*h)^5.25588   # (Pa)
# ds$AirTemp <- 25 #(degrees Celsius)

ds <- ds[!duplicated(ds$TIMESTAMP),] # extract duplicated rows

# interpolate measurement timestamps for whole dataset
##create  continuous timestamp vector (here: interval=1 second)
regular_timesteps <- seq(min(ds$TIMESTAMP),max(ds$TIMESTAMP), by="1 sec")

collar_df <- tibble("TIMESTAMP"=ds$TIMESTAMP,"collar"=ds$solenoid_valvesInt)

ds <-convertToRegularTimesteps(ds,c("CO2_dry", "CH4_dry","H2Oppt", "NH3_dry","N2O_dry"),regular_timesteps)
ds <- left_join(ds,collar_df,by=join_by("TIMESTAMP"=="TIMESTAMP")) %>% fill(.,collar,.direction="down")

nrow(ds_subset)
nrow(ds)

#-- In order to process each measurement cycle independently,
#-- we first determine parts of the time series that are contiguous,
#-- i.e. without gaps and without change of an index variable, here variable collar.


ds$TIMESTAMP_hour <- floor_date(ds$TIMESTAMP, unit = "hour") #create a column with a timestamp roundet to the hour, in order to join with hourly evironmental data

#join Temp_PressureData with ds  (approx does not work for more following NA's, just fill nas with same value as last one (function "fill" from tidyr package))
ds <- left_join(ds,Additional_Weather_Data ,by=join_by("TIMESTAMP_hour"=="DATETIME_hourly")) %>% fill(.,h,AirTemp,Pa,rel_humidity,shortwave_radiation,ET0, .direction = "up")

ds$VPD <- calcVPD( ds$AirTemp, ds$Pa, ds$H2Oppt) ## Here we calculate plant-to-air vapour pressure deficit
```

# Data overview

Plots of the whole time series give a good overview about the data.

```{r}
p_collar <- plot(ds$TIMESTAMP,ds$collar, pch = ".", ylab = "collar (Chamber)",xlab = "Time")

### facet plot of time series (for whole subset)
ds_gas_long <- gather(ds, key="gas", value="value", c("CO2_dry","H2Oppt","CH4_dry","NH3_dry","N2O_dry"))
p_gas_facet <- ggplot(ds_gas_long, aes(x=TIMESTAMP, y=value))+
  ggtitle(format(ds$TIMESTAMP,"%d/%m/%Y")[1])+
  geom_point(pch = ".")+
  facet_wrap(~factor(gas,levels=c("CO2_dry","H2Oppt","CH4_dry","NH3_dry","N2O_dry")),ncol=1,scales = "free")

ds_envar_long <- gather(ds, key="envar", value="value", c("AirTemp","Pa","rel_humidity","shortwave_radiation","ET0"))
p_envar_facet <-  ggplot(ds_envar_long, aes(x=TIMESTAMP, y=value))+
  ggtitle(format(ds$TIMESTAMP,"%d/%m/%Y")[1])+
  geom_point(pch = ".")+
  facet_wrap(~factor(envar,levels=c("AirTemp","Pa","rel_humidity","shortwave_radiation","ET0")),ncol=1,scales = "free")

p_collar
p_gas_facet
p_envar_facet
```

# Chunk creation
In order to process each measurement cycle independently, we first determine parts of the time series that are contiguous, i.e. without gaps and without change of an index variable, here variable collar.

Too many data points may slow down calculations, therefore, data can be thinned. The thinning interval can be set with 'slice(seq(1,n(),[THINNINGINTERVAL]))'. The object 'mapped_collars' shows which collar is assigned to which chunk.

For each collar (chamber), specific settings can be defined in the tibble "collar_spec".


```{r chunks}
dsChunk_raw <- subsetContiguous(ds, colTime = "TIMESTAMP", colIndex = "collar",
                            gapLength = 12, minNRec = 180, minTime = 180, indexNA = 13)

# thin data: select the thinning interval to make calculations more efficient
dsChunk <- dsChunk_raw %>% group_by(iChunk) %>% slice(seq(1, n(), 8)) %>% ungroup()

mapped_collars <- dsChunk %>% group_by(iChunk) %>% summarise(collar = first(collar)) %>%  head()

## DataFrame collar_spec then needs to specify for each collar id in column collar,
# the colums area (m2) and volume (m3), as well a tlag (s), the lag time between start of the cycle , i.e. the start of the chunk (usually chamber closing time), and the time when the gas reaches the sensor.

collar_spec <- tibble(
  collar = unique(dsChunk$collar),
  depth = rnorm(9,0.1,0.01), #here the depth of each chamber needs to be defined
  area = surfaceArea,
  volume = chamberVol + surfaceArea * depth,
  tlag = NA)
head(collar_spec)

```

# Plots of individual chunks
Time series of individual chunks can be plotted with the function 'chunk_plot()'.

```{r}
# Generate labels for each gas
labels_CO2 <- chunk_labels(dsChunk, CO2_dry, 1.05)
labels_H2O <- chunk_labels(dsChunk, H2Oppt, 1.05)
labels_CH4 <- chunk_labels(dsChunk, CH4_dry, 1.05)
labels_N2O <- chunk_labels(dsChunk, N2O_dry, 1.05)
labels_NH3 <- chunk_labels(dsChunk, NH3_dry, 1.05)

# Generate plots for each gas
p_CO2 <- chunk_plot(dsChunk, CO2_dry, labels_CO2)
p_H2O <- chunk_plot(dsChunk, H2Oppt, labels_H2O)
p_CH4 <- chunk_plot(dsChunk, CH4_dry, labels_CH4)
p_N2O <- chunk_plot(dsChunk, N2O_dry, labels_N2O)
p_NH3 <- chunk_plot(dsChunk, NH3_dry, labels_NH3)

##save the plots
# for (p in c("p_CO2","p_H2O","p_CH4","p_N2O","p_NH3")) {
#   ggsave(filename=paste0(results_dir,"/",str_sub(fileName,end=-18),"_allchunks_",p,".pdf"),get(p),width = 40,height = 30,units = "cm")
# }

p_CO2
```

# Calculation of the gas fluxes for multiple chunks
```{r}
#One can save processing time and avoid failures in the non-robust breakpoint-detection by specifying a fixed lag-time (may differ across collars) with the collar specification.
collar_spec_CO2 <-mutate(collar_spec, tlag = 60,depth = 0.1, volume =  chamberVol + surfaceArea * depth )  # volume has to be re-calculated here otherwise only depth column will change if we mutate depth
collar_spec_H2O <-mutate(collar_spec, tlag = 120)
collar_spec_CH4 <-mutate(collar_spec, tlag = 60)
collar_spec_NH3 <-mutate(collar_spec, tlag = 60)
collar_spec_N2O <-mutate(collar_spec, tlag = 120)

# parallel computing
plan(multisession, workers = 4) 

res_CO2 <- calcClosedChamberFluxForChunkSpecs(
  dsChunk, collar_spec_CO2
  , colTemp = "AirTemp", colPressure = "Pa"
  , fRegress = c(lin = regressFluxLinear, tanh = regressFluxTanh, exp = regressFluxExp, poly= regressFluxSquare)
  , debugInfo = list(omitEstimateLeverage = FALSE)	# faster
  , colConc = "CO2_dry", colTime = "TIMESTAMP"
  , concSensitivity = 0.01
)


res_NH3 <- calcClosedChamberFluxForChunkSpecs(
  dsChunk, collar_spec_NH3
  , colTemp = "AirTemp", colPressure = "Pa"
  , fRegress = c(lin = regressFluxLinear, tanh = regressFluxTanh, exp = regressFluxExp, poly= regressFluxSquare)
  , debugInfo = list(omitEstimateLeverage = FALSE)	# faster
  , colConc = "NH3_dry", colTime = "TIMESTAMP"	# colum names conc ~ timeInSeconds
  , concSensitivity = 0.1
)

res_CH4 <- calcClosedChamberFluxForChunkSpecs(
  dsChunk, collar_spec_CH4
  , colTemp = "AirTemp", colPressure = "Pa"
  , fRegress = c(lin = regressFluxLinear, tanh = regressFluxTanh, exp = regressFluxExp, poly= regressFluxSquare)
  , debugInfo = list(omitEstimateLeverage = FALSE)	# faster
  , colConc = "CH4_dry", colTime = "TIMESTAMP"	# colum names conc ~ timeInSeconds
  , concSensitivity = 0.01
)

res_N2O <- calcClosedChamberFluxForChunkSpecs(
  dsChunk, collar_spec_N2O
  , colTemp = "AirTemp", colPressure = "Pa"
  , fRegress = c(lin = regressFluxLinear, tanh = regressFluxTanh, exp = regressFluxExp, poly= regressFluxSquare)
  , debugInfo = list(omitEstimateLeverage = FALSE)	# faster
  , colConc = "N2O_dry", colTime = "TIMESTAMP"	# colum names conc ~ timeInSeconds
  , concSensitivity = 0.01
)

res_H2O <- calcClosedChamberFluxForChunkSpecs(
  dsChunk, collar_spec_H2O
  , colTemp = "AirTemp", colPressure = "Pa"
  , fRegress = c(lin = regressFluxLinear, tanh = regressFluxTanh, exp = regressFluxExp, poly= regressFluxSquare)
  , debugInfo = list(omitEstimateLeverage = FALSE)	# faster
  , colConc = "H2Oppt", colTime = "TIMESTAMP"	# colum names conc ~ timeInSeconds
  , concSensitivity = 0.01
)

```


# Inspection of the results
The results can be inspected:
```{r}
res_CO2
res_CH4
res_N2O
res_NH3
res_H2O

```



# Visualisation of the results
The results can be visualized in a plot:
```{r}
#Plot the results including the fits as facet plots
res_facets_CO2 <- plotCampaignConcSeries(ds=dsChunk,varName = "CO2_dry",dsFits = res_CO2, plotsPerPage = 64L,fileName =paste0(results_dir,"/CO2_fit_facets.pdf")) #fileName =paste0(results_dir,"/CO2_fit_facets.pdf") )
print( res_facets_CO2$plot[[1]])


res_facets_NH3 <- plotCampaignConcSeries(ds=dsChunk,varName = "NH3_dry",dsFits = res_NH3, plotsPerPage = 64L,fileName =paste0(results_dir,"/NH3_fit_facets.pdf")) #fileName =paste0(results_dir,"/CO2_fit_facets.pdf") )
print( res_facets_NH3$plot[[1]])


res_facets_CH4 <- plotCampaignConcSeries(ds=dsChunk,varName = "CH4_dry",dsFits = res_CH4, plotsPerPage = 64L,fileName =paste0(results_dir,"/CH4_fit_facets.pdf")) #fileName =paste0(results_dir,"/CO2_fit_facets.pdf") )
print( res_facets_CH4$plot[[1]])

res_facets_N2O <- plotCampaignConcSeries(ds=dsChunk,varName = "N2O_dry",dsFits = res_N2O, plotsPerPage = 64L,fileName =paste0(results_dir,"/N2O_fit_facets.pdf")) #fileName =paste0(results_dir,"/CO2_fit_facets.pdf") )
print( res_facets_N2O$plot[[1]])


res_facets_H2O <- plotCampaignConcSeries(ds=dsChunk,varName = "H2Oppt",dsFits = res_H2O, plotsPerPage = 64L,fileName =paste0(results_dir,"/H2O_fit_facets.pdf")) #fileName =paste0(results_dir,"/CO2_fit_facets.pdf") )
print( res_facets_H2O$plot[[1]])

```

The fitted form (solid line) approximates the change of measured concentration (points) over time. Usually there is a lag-time (dotted vertical line) between concentration change, and when the time when the signal reaches the sensor. The flux after the intial lag-time is reported at the top left together with its uncertainty in mumol/m2/s. The label can be provided to distinguish different measurement cycles.



