---
title: "ChamberProcWorkflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ChamberProcWorkflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
# load libraries
library(ChamberProc)
library(ChamberProc)
library(RespChamberProc)
library(rlang)
library(changepoint)
library(nlme)
library(segmented)
library(tibble)
library(plyr)
library(dplyr)
library(purrr)
library(stringr)
library(ggplot2)
library(data.table)
library(lubridate)
library(tidyr)
library(doSNOW)
library(furrr)
library(elevatr)
library(sf)
library(openmeteo) #retrieves historical weather data from https://open-meteo.com/en/docs/historical-weather-api

```

```{r user-defined input}
#set working directory
#setwd("/Users/ms/Research Projects/Espana/UCO_Chamber/1_Test_newPackage_ChamberProc")

latDeciDeg <- 37.91514875822371 #enter here latitude in Geographical Coordinates (decimal degrees); crs =4326)
lonDeciDeg <- -4.72405536319233 #enter here longitude in Decimal Degrees


```

```{r load-data}
ds0 <- sample_PICARRO_data

ds0$TIMESTAMP <- as.POSIXct(paste0(ds0$DATE," ",ds0$TIME), "%Y-%m-%d %H:%M:%S", tz= "UTC")

# Subset data
ds_subset <- subset(ds0, as.numeric(TIMESTAMP) >= as.numeric(RespChamberProc::as.POSIXctUTC("2022-05-31 08:00:00")) )
ds_subset <- subset(ds_subset, as.numeric(TIMESTAMP) <= as.numeric(RespChamberProc::as.POSIXctUTC("2022-05-31 09:30:00" )) )

ds <- ds_subset

```

# Correct Gases

The PICARRO IRGA gives dry mole fractions for CO2, N2O, CH4, but not for NH3 and H2O. Therefore these gases have to be corrected. Additional envionmental data (h, Pa and AirTemp are crucial) is loaded with the function "Additional_Weather_Data". Alternatively, values for `r ds$h` and `r ds$AirTemp` ,and `r ds$AirTemp` can be assigned manually.

```{r}
Additional_Weather_Data <- getAdditionalWeatherVariables(latDeciDeg, lonDeciDeg,format(min(ds$TIMESTAMP),"%Y-%m-%d"),format(max(ds$TIMESTAMP),"%Y-%m-%d"))


ds$collar<-  ds$solenoid_valves %>% as.integer()
ds$H2Oppt <- ds$H2O*10 # H2O from PICARRO is in %. Needs to be in ppt --> We need to multiply by 10
ds$N2O_dry <- ds$N2O_dry1min
ds$NH3_dry <- 10E-3*corrConcDilution(ds, colConc = "NH3", colVapour = "H2Oppt")  #NH3 from PICARRO is in ppb --> multiply colVapour by 10^-3 to get ppm
ds$H2O_dry <- corrConcDilution(ds, colConc = "H2Oppt", colVapour = "H2Oppt")


ds <- ds[!duplicated(ds$TIMESTAMP),] # extract duplicated rows


ds$TIMESTAMP_hour <- floor_date(ds$TIMESTAMP, unit = "hour") #create a column with a timestamp roundet to the hour, in order to join with hourly evironmental data

#join Temp_PressureData with ds  (approx does not work for more following NA's, just fill nas with same value as last one (function "fill" from tidyr package))
ds <- left_join(ds,Additional_Weather_Data ,by=join_by("TIMESTAMP_hour"=="DATETIME_hourly")) %>% fill(.,h,AirTemp,Pa,rel_humidity,shortwave_radiation,ET0, .direction = "up")

ds$VPD <- calcVPD( ds$AirTemp, ds$Pa, ds$H2Oppt) ## Here we calculate plant-to-air vapour pressure deficit

```

# Data overview

Plots of the whole time series give a good overview about the data.

```{r}
p_collar <- plot(ds$TIMESTAMP,ds$collar, pch = ".", ylab = "collar (Chamber)",xlab = "Time")

ds_gas_long <- gather(ds, key="gas", value="value", c("CO2_dry","H2Oppt","CH4_dry","NH3_dry","N2O_dry"))
p_gas_facet <- ggplot(ds_gas_long, aes(x=TIMESTAMP, y=value))+
  ggtitle(format(ds$TIMESTAMP,"%d/%m/%Y")[1])+
  geom_point(pch = ".")+
  facet_wrap(~factor(gas,levels=c("CO2_dry","H2Oppt","CH4_dry","NH3_dry","N2O_dry")),ncol=1,scales = "free")

ds_envar_long <- gather(ds, key="envar", value="value", c("AirTemp","Pa","rel_humidity","shortwave_radiation","ET0"))
p_envar_facet <-  ggplot(ds_envar_long, aes(x=TIMESTAMP, y=value))+
  ggtitle(format(ds$TIMESTAMP,"%d/%m/%Y")[1])+
  geom_point(pch = ".")+
  facet_wrap(~factor(envar,levels=c("AirTemp","Pa","rel_humidity","shortwave_radiation","ET0")),ncol=1,scales = "free")

p_collar
p_gas_facet
p_envar_facet

```

# Chunk createion
In order to process each measurement cycle independently, we first determine parts of the time series that are contiguous, i.e. without gaps and without change of an index variable, here variable collar. Measurement data can have non-equidistant timesteps. Too many data points may slow down calculations, therefore, data can be thinned. For each collar (chamber), specific settings can be defined in the tibble "collar_spec".

```{r chunks}
dsChunk_raw <- subsetContiguous(ds, colTime = "TIMESTAMP", colIndex = "collar",
                            gapLength = 12, minNRec = 180, minTime = 180, indexNA = 13)

# thin data: select the thinning interval to make calculations more efficient
dsChunk <- dsChunk_raw %>% group_by(iChunk) %>% slice(seq(1, n(), 8)) %>% ungroup()

mapped_collars <- dsChunk %>% group_by(iChunk) %>% summarise(collar = first(collar)) %>%  head()



## DataFrame collar_spec then needs to specify for each collar id in column collar,
# the colums area (m2) and volume (m3), as well a tlag (s), the lag time between start of the cycle , i.e. the start of the chunk (usually chamber closing time), and the time when the gas reaches the sensor.

chamberVol = 0.6*0.6*0.6		# chamber was a cube of 0.6m length
surfaceArea = 0.6*0.6

collar_spec <- tibble(
  collar = unique(dsChunk$collar),
  depth = pmax(0,rnorm(length(collar), mean = 0.03, sd = 0.015)),
  area = surfaceArea,
  volume = chamberVol + surfaceArea * depth,
  tlag = NA)
head(collar_spec)

```
